---
title: "Authorship verification using the package `stylo`"
author: Maciej Eder
date: 26/05/2018
output: 
  html_document:
    highlight: pygments
    fig_width: 7
    fig_height: 5
bibliography: bibliography_subset.bib
csl: apa.csl
---



```{r setup, include = FALSE}
# set global chunk options

library(stylo)
library(knitr)

# using a cache folder
opts_chunk$set(cache = TRUE)
```



## Introduction

This post introduces to a new feature of the package `stylo` (ver. 0.6.7), namely the General Imposters (GI) method, also referred to as the second verification system (o2), introduced by Koppel and Winter [-@koppel_determining_2014] and applied to the study of Julius Caesar’s disputed writings [@kestemont_authenticating_2016]. To quote the authors, “[t]he general intuition behind the GI, is not to assess whether two documents are simply similar in writing style, given a static feature vocabulary, but rather, it aims to assess whether two documents are significantly more similar to one another than other documents, across a variety of stochastically impaired feature spaces [@eder_computational_2012; @stamatatos_authorship_2006], and com- pared to random selections of so-called distractor authors [@juola_rowling_2015], also called ‘imposters’.” [@kestemont_authenticating_2016, 88].

The implementation provided by the package `stylo` is a rather faithful interpretation of two algorithms described in a study on authorship verification [@kestemont_authorship_2016]. A tiny function for computing the score `c@1` used to evaluate the system is directly transplanted from the original implementation ([https://github.com/mikekestemont/ruzicka](https://github.com/mikekestemont/ruzicka)).

The main procedure is available via the function `imposters()`. It assumes that all the texts to be analyzed are already pre-processed and represented in a form of a matrix with frequencies of features (usually words). The function contrasts, in several iterations, a text in question against (1) some texts written by possible candidates to authorship, or the authors that are suspected of being the actual author, and (2) a selection of “imposters”, or the authors that could not have written the text to be assessed. Consequently, a given candidate’s class is assigned a score between 0 and 1. 

On theoretical grounds, any score above 0.5 would suggest that the authorship verification for a given candidate was successful. However, the scores such as 0.39 or 0.63 should be considered suspicious: they seem to suggest that the classifier had problems in making clear-cut decisions. Another function, namely `imposters.optimize()`, is meant to assess – via a grid search – optimal parameters defining the above-mentioned grey area where the classifier should keep shut. It is a procedure of testing iteratively each text from a training corpus (one at a time) against all the possible candidates. Being computationally intense, this is a rather time-consuming task: be prepared to leave you machine running for a few hours.






## Installation

The latest (and stable) version of the package `stylo` is usually available on CRAN a few days after such a new version is released. In this case, the installation is trivial:

``` {r eval = FALSE}
install.pacgages("stylo")
```

Maybe a better way to install brand-new and/or experimental versions of the package is to grab it directly from the GitHub repository – please make sure, however, that you have the package `devtools` installed in your system. The next step is straightforward:

``` {r eval = FALSE}
library(devtools)
install_github("computationalstylistics/stylo")
```

If no errors occurred during the installation, we’re all set!



## A tl;dr working example

To test at a glance what the `imposters()` function can offer, type the following code. Notice that you don’t even need to prepare any corpus beforehand:

``` {r eval = FALSE}
# activating the package 'stylo':
library(stylo)

# activating one of the datasets provided by the package 'stylo';
# this is a table of frequences of a few novels, including "The Cuckoo's Calling"
# by Robert Galbraith, aka JK Rowling:
data(galbraith)

# to learn more about the dataset, type:
help(galbraith)

# to see the table itself, type:
galbraith

# now, time for launching the imposters method:
imposters(galbraith)
```

After a few seconds, the final results will be shown on the screen:

``` {r echo = FALSE}
library(stylo)
data(galbraith)
imposters(galbraith)
```

The interpretation of the results is rather straightforward: “The Cuckoo's Calling” was not written by CS Lewis, nor was it penned by JRR Tolkien. Similarly, the score for JK Rowling turned out to be very high (could not be higher), which clearly suggests the author of the analyzed novel. The only class that is difficult to interpret is Harlan Coben, since he was assigned 0.29. The score obviously falls into the <0.5 category, but still: can we reliably say that this candidate should be rulled out or, rather, should we abstain from any conclusions here? The problem of defining the grey area around the score 0.5 will be discussed below.




## Details

As you might have noticed, the dataset `galbraith` contains frequencies for different texts by a few authors, the class _GALBRAITH_ being represented by a single text, though. Having no additional parameters, the function tries to identify such a single text and then assumes that this is the anonymous sample to be assessed. 

In a vast majority of cases, however, one would like to have some control on choosing the text to be contrasted against the corpus. The function provides a dedicated parameter `test` to do the trick. Note the following code:

``` {r eval = FALSE}
# getting the 8th row from the dataset (it contains frequencies for Galbraith):
my_text_to_be_tested = galbraith[8,]

# building the reference set so that it does not contain the 8th row
my_frequency_table = galbraith[-c(8),]

# launching the imposters method:
imposters(reference.set = my_frequency_table, test = my_text_to_be_tested)
```

Consequently, if you want to test who wrote “The Lord of the Rings”, part 1, get the 24th row from the table: 

``` {r eval = FALSE}
my_text_to_be_tested = galbraith[24,]
my_frequency_table = galbraith[-c(24),]
imposters(reference.set = my_frequency_table, test = my_text_to_be_tested)
```

So far, I’ve been neglecting one important feature of the imposters method. As Kestemont et al. [-@kestemont_authorship_2016] show in their “Algorithm 1”, the method tries to compare an anonymous text against (1) a candidate set, containing the works of a probable candidate author, and (2) the imposters set, containing lots of text by people who could not have written the text in question. In the previous examples, where no canditate set was explicitly indicated, the method repeatedly tested all of possible authors as potential candidates (one at a time). It is a time consuming task. If you plan to focus on just one _actual_ candidate, e.g. if you want to test if “The cuckoo’s Calling” was written by JK Rowling, you should define the parameters as follows:

``` {r eval = FALSE}
# indicating the text to be tested (here, "The cuckoo's Calling"):
my_text_to_be_tested = galbraith[8,]

# defining the texts by the candidate author (here, the texts by JK Rowling):
my_candidate = galbraith[16:23,]

# building the reference set by excluding the already-selected rows
my_imposters = galbraith[-c(8, 16:23),]

# launching the imposters method:
imposters(reference.set = my_imposters, test = my_text_to_be_tested, candidate.set = my_candidate)
```

The above code shows a standard application of the General Imposters method. In practice, however, I’d rather test _all_ the authors iteratively, even if this requires quite a lot of time to complete the task. This will serve as an additional cross-validation step, because I will be able to check if all the remaining candidates will be assigned zeros (as expected). In other words, I’d rather compare the behavior of JK Rowling in comparison to all possible candidate authors, even if Rowling is the only class I believe would get a reasonable score.






## Loading a corpus from files

I am fully aware that the function `imposters()` in its current form requires some advanced knowledge of R, since it does not provide any corpus pre-processing (as the main functions of the package `stylo` do). Specifically, one needs to know in advance how to produce a table of frequencies. This step has been already described elsewhere [@eder_stylometry_2016, 109–111], therefore I will not go into nuanced details here. A straightforward way to get from raw text files to the imposters results might look as follows:

``` {r eval = FALSE}
# activating the package
library(stylo)

# setting a working directory that contains the corpus, e.g.
setwd("/Users/m/Desktop/A_Small_Collection_of_British_Fiction/corpus")

# loading the files from a specified directory:
tokenized.texts = load.corpus.and.parse(files = "all")

# computing a list of most frequent words (trimmed to top 2000 items):
features = make.frequency.list(tokenized.texts, head = 2000)

# producing a table of relative frequencies:
data = make.table.of.frequencies(tokenized.texts, features, relative = TRUE)

# who wrote "Pride and Prejudice"? (in my case, this is the 4th row in the table):
imposters(reference.set = data[-c(4),], test = data[4,])
```





## Optimizing the decision scores


Score shifting algorighm

```
imposters.optimize()
```

Be prepared to wait...



## Parameters

In its current form, the function `imposters()` works with the Delta method only. Next versions will provide SVM, NSC, kNN and NaiveBayes. As most of you know very well, the general Delta framework can be combined with many different distance measures. E.g. in their paper introducing the imposters method [@kestemont_authorship_2016], the authors argue that the Ruzicka metrics (aka Minmax) outperforms other measures. Similarly, the Wurzburg guys [@evert_understanding_2017] show that Cosine Delta rocks when compared to other distances. My implementation of the `imposters()` applies Classic Delta by default, but other measures can be used as well. Try the following options:


``` {r eval = FALSE}
# activating the package 'stylo':
library(stylo)

# activating one of the datasets provided by the package 'stylo':
data(galbraith)

# Classic Delta distance
imposters(galbraith, distance = "delta")

# Cosine Delta (aka Wurzburg Distance)
imposters(galbraith, distance = "wurzburg")

# Ruzicka Distance (aka Minmax Distance)
# (please keep in mind that it takes AGES to compute it!)
imposters(galbraith, distance = "minmax")
```

Not really impressed, right? This is because the signal of JK Rowling is really strong, and all the measures perform just fine. Let’s try something more difficult. Did you know that “In Cold Blood” by Truman Capote is stylometrically hard to associate with its actual author? Execute the following code:

``` {r eval = FALSE}
# activating the package 'stylo':
library(stylo)

# activating another dataset, which contains Southern American novels:
data(lee)

# defining the test text, i.e. "In Cold Blood"
my_text_to_be_tested = lee[1,]

# defining the comparison corpus
my_reference_set = lee[-c(1),]

# Classic Delta distance
imposters(my_reference_set, my_text_to_be_tested, distance = "delta")

# Eder's Delta distance
imposters(my_reference_set, my_text_to_be_tested, distance = "eder")

# Cosine Delta (aka Wurzburg Distance)
imposters(my_reference_set, my_text_to_be_tested, distance = "wurzburg")

# Ruzicka Distance (aka Minmax Distance)
# (please keep in mind that it takes AGES to compute it!)
imposters(my_reference_set, my_text_to_be_tested, distance = "minmax")
```

Have you noticed the amazing improvement of Wurzburg Delta over the other measures? It’s really cool!

Certainly, the same applied to the `imposters.optimize()` function. The same parameters that were introduced immediately above, can be passed to the fine-tuning function, e.g.:

``` {r eval = FALSE}
results = imposters(my_reference_set, distance = "wurzburg")
```


## References

